---
title: Function documentation
author: Abhijit Dasgupta
date: '2018-11-16'
slug: function-documentation
draft: false
categories:
  - R
tags:
  - Good Practices
menu: ''
---

## Writing functions

We learned how to write functions in class this week. As a reminder, a function takes a
particular structure, starting with the keyword `function`. In RStudio, if you type `fun [TAB]` (i.e., type `fun` and then hit the tab key), you get the following function scaffolding:

```{r, eval = F}
name <- function(variables) {
  
}
```

`name` is the name of the function, the inputs/arguments go inside the `()`, and the lines of 
code that comprise the function go between the `{}` in separate lines. A simplistic function might be to get the mean and median of a vector of numbers:
```{r}
summaries <- function(x) {
  mn <- mean(x, na.rm=T)
  md <- median(x, na.rm=T)
  return(tibble('mean' = mn, 'median' = md))
}
```

> I always recommend putting whatever the function returns within the `return` function, so there is no confusion about what the results/output of the function is

## Documenting your function

In a simple function like this, it is straightforward to understand what is going on. But you'll soon be writing more complex functions, where you need the input(s) to be of a certain type and the resulting output to be of some type, and you also want to remind yourself in 3 months or a year what this function was written for. There is a very nice package called `roxygen2` that allows the documentation to stay with the function. Though the package was built to help document functions that are part of R packages, it can be useful for _ad hoc_ functions within your analysis project as well. And best of all, it is baked in to RStudio!.

To create the documentation scaffolding, put your cursor on the first line of the function, and then either type `Ctrl-Alt-Shift-R` (Win/Linux) or `Cmd-Alt-Shift-R` (Mac) or, from the menu, `Code > Insert Roxygen Skeleton`. This is what that scaffolding looks like:

```{r}
#' Title
#'
#' @param x 
#'
#' @return
#' @export
#'
#' @examples
summaries <- function(x) {
  mn <- mean(x, na.rm=T)
  md <- median(x, na.rm=T)
  return(tibble('mean' = mn, 'median' = md))
}
```

The `roxygen` documentation lines start with  `#'`, which R treats like a line of comments and ignores. The scaffolding provides a particular format to the documentation. The first line should be the title of the documentation, then you can skip one line (keeping the `#'` on that line) and then write paragraphs describing the function. The `@param` fields are auto-filled with the explicit arguments in your function, and you can then describe what those arguments mean. The `@return` field describes what the result of your function is. The `@export` tag is useful when you're building packages, and indicates that the function will be part of the main package functions. Lastly, `@examples` asks you to optionally provide an example of how the function works. So, for the example I'm using here, we might fill out the documentation like this: 

```{r}
#' Summarizing data
#' 
#' This function summarizes the data in a numerical vector by computing the mean and median of
#' the data
#'
#' @param x A numeric vector
#'
#' @return A tibble that gives the mean and median of the data in `x`, 
#' @export
#'
#' @examples
#' summaries(mtcars$mpg)
summaries <- function(x) {
  mn <- mean(x, na.rm=T)
  md <- median(x, na.rm=T)
  return(tibble('mean' = mn, 'median' = md))
}
```

More detail about roxygen, and it's role in documenting functions within packages, are available from Karl Broman's excellent [primer](https://kbroman.org/pkg_primer/pages/docs.html){target="_blank"}
